---
title: "Alex Project2 Code Testing"
author: "Xianming Tan"
date: "2024-06-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

To understand Alex's code for project 2. 

See https://github.com/aquinter96/Dissertation_Projects/tree/main/Project_2/Initial_Estimate_Comparisons_Code.  

We first work on `Truth_Inits.R`. 

**Completed**

- wrap into several functions
  - model parameters setup ... 
    - `Model_Meta_Param.R` (dimensions)
    - `Model_Matrix_Param.R` (matrices like A1/A2/...)
  - data generation
    - `Data_Generation.R` (a function)
  - estimation 
    - `_EMAlgXAdLassoCV_00.R`
    - `_EMAlgYAdLassoCV_00.R`
  - testing data and results 
    - `TestData.RData`
  - to re-organize: `_EMAlgYAdLassoCV_00.R` 
    - See `EMAlgYAdLassoCV_01.R`
    - Main function `EMAlgYAdLasso()`
    - The E-step part: EMAlgYAdLasso_EStep() 
    - The M-step part: EMAlgYAdLasso_MStep()

**To dos**

  - estimation functions are too long, will divide it into multiple parts 
    - `_EMAlgXAdLassoCV_00.R` (X-part ... )

## Setup 

### Load libraries

```{r, echo=TRUE}
library(MASS)
library(pracma)
library(psych)
library(r.jive)
```

### Random seed 

```{r, echo=TRUE}
## args <- commandArgs(TRUE)
# datset <- 1 ## as.numeric(args[1])
set.seed(12543)
```

### To initialize parameters

The model for Paper 2 is described in overleaf and in `Paper2 overview.md` (obsidian note)

In the below code, we setup the parameters for the model

**Some comments**

- We used `rnorm()` to generate elements for A1/A2/B1/B2/Phi etc 
- This is not ideal, especially for the residual covariance part, because this will create some very difficult scenario in factor analysis. 
- For residual covariance, it's better to set nearly identical values for the elements. 
- This, however, is easy to fix. 
- We move the following part to 2 separate R files.
  - Model_Meta_Param.R (for )
  - Model_Matrix_Param.R (For A1/A2/B1/B2/...)
    - For this part, we may replace `rnorm` by `runif`

```{r, echo=TRUE}
code_dir <- "C:/Users/tanxi/Downloads/Dissertation_Projects-main/Project_2/"
source(file = paste0(code_dir, "Model_Meta_Param.R") )
source(file = paste0(code_dir, "Model_Matrix_Param.R") )
```

### To generate simulated data 

- Sample size: $n = 1000$.
- In this part, we generate $X,Y$
- We wrap this part to a R function `Data_Generation` and saved in  file `Data_Generation.R`


```{r, echo=TRUE}
source(file = paste0(code_dir, "Data_Generation.R") )

set.seed(123456)
MyData <- Data_Generation(n = 1000, Model_Params=Model_Params)
```

### True values of G1-G4, for intermediate checking ... 

```{r, echo=TRUE}
G1t <- Model_Params$Gamma1%*%Model_Params$Theta + Model_Params$Delta1%*%Model_Params$Pi1
G2t <- Model_Params$Gamma2%*%Model_Params$Theta + Model_Params$Delta2%*%Model_Params$Pi2
G3t <- Model_Params$Gamma1%*%Model_Params$Psi + Model_Params$Delta1%*%Model_Params$Omega1
G4t <- Model_Params$Gamma2%*%Model_Params$Psi + Model_Params$Delta2%*%Model_Params$Omega2
Gt <- rbind(cbind(G1t, G3t), cbind(G2t, G4t))

D <- cov(cbind(t( MyData$X1), t(MyData$X2)))
```

### To get initial values on A/B matrices for the X part ... 

- In this version, we used the true parameter values as the initial values for EM algorithm

## EM algorithms  

These parameters are:  $A_1, A_2,  B_1, B_2, \Phi_R, \Phi_{S1}, \Phi_{S2}$

The following code chunk defines a function `EMAlgXAdLassoCV` to estimate these parameters given data $X$ and initial parameters. 

We first move function `EMAlgXAdLassoCV` to R file "_EMAlgXAdLassoCV_00.R" (this will not replaced by other files not starting with "_"). 


### To test the EM algorithm for the X part . 

```{r, echo=TRUE}
source(file = paste0(code_dir, "_EMAlgXAdLassoCV_00.R") )
```

To do:  to re-organize the EM algorithm for the X-part.  

```{r, echo=TRUE}
time0 <- proc.time()
############################################################################################################################################## 
## This test fix tuning parameter at 0. This means, this compuation is really intensive and slow (8 minutes for one X data). This is weird. 
############################################################################################################################################## 
test <- EMAlgXAdLassoCV(Xpanel = list(t(MyData$X1), t(MyData$X2)), x_joint =  meta_param$m, x_indiv = c(meta_param$u1, meta_param$u2), 
                        Ainits = list(Model_Params$A1, Model_Params$A2), Binits = list(Model_Params$B1, Model_Params$B2), 
                        vcovinits = list(Model_Params$PhiR, Model_Params$PhiS1, Model_Params$PhiS2, Model_Params$Phi11, Model_Params$Phi12), 
                        tuningpA = 0, tuningpB = 0,
                        weightsA = list(Model_Params$A1, Model_Params$A2), weightsB = list(Model_Params$B1, Model_Params$B2)  )
time1 <- proc.time()
timeDiff <- time1 - time0 ## 8 minutes for one data set (X part only)  ...; 150s (2.5 minutes)  

timeDiff 
```

### To test the EM step for the Y part

- The initials from the ADAM max sometimes not good, and hence EM did not improve much ...  
- I think that we used the true parameter values as the initial value for EM algorithm, right?

```{r, echo=TRUE}
# source(file = paste0(code_dir, "_EMAlgYAdLassoCV_00.R") )
source(file = paste0(code_dir, "EMAlgYAdLassoCV_01.R"))
```

- The Y-part is re-organized into several functions ... 


```{r, echo=TRUE}
time0 <- proc.time()
## fixed tuningpGamma,  tuningpDelta
## using true A1/A2/B1/B2 as initials ... 
## Example:  
myEMRes <-       EMAlgYAdLasso(Data = MyData,  
                              initial_Model_Param = Model_Params, 
                              tuningpGamma = 0, 
                              tuningpDelta = 0, 
                              weightsGamma = list(Model_Params$Gamma1, Model_Params$Gamma2), 
                              weightsDelta = list(Model_Params$Delta1, Model_Params$Delta2) )

time1 <- proc.time()
timeDiff <- time1 - time0    

timeDiff

plot(myEMRes$log.Like, main = "Change in Log-likelihood", xlab = "Iterations", ylab = "log-likelihood")

plot(myEMRes$diffList, main = "Change in max abs diff in parameters", xlab = "Iterations", ylab = "max-abs-diff")

# save(test, test2, meta_param, Model_Params, MyData, file=paste0(code_dir,  "TestData.RData"))
``` 
